Safe Memory Regions for Big Data Processing

### 2016-06-01 paper/type-inference.tex
Type inference description

Fig.~\ref{fig:type-inference-algo} presents the high-level outline of the type
inference algorithm.
The algorithm consists of the following steps:
\begin{enumerate}
 \item The first step elaborates the input program by introducing \emph{(formal and actual) region parameters},
and \emph{region variables} (representing yet undetermined \emph{actual region parameters})
as follows:
\begin{enumerate}
\item For every class determine a set of \emph{region parameters} ($\rho$)
       it should be parametrized over.
\item For every field of the class, determine its region parametrized type by
       generating as many fresh \emph{actual region parameters} as required by its class (type),
       which in turn become formal region parameters of the containing class.
\item  For every method and function type, determine its corresponding region
   type template by introducing fresh region parameters as required by
   the type of each parameter and return value (in addition to the allocation region
   parameter).
 \item Elaborate every \C{new} expression, method call or function application by
   introducing \emph{fresh region variables} (as required by the region type
   template of the called method or function).
\item Introduce for each class and method, a \emph{predicate variable} ($\varphi$) to denote
   an undetermined set of constraints over the region parameters of that
   class/method.
\end{enumerate}
 \item In the second step, we analyze the program to generate a set of constraints that capture
   well-formedness requirements and subtyping relationships between type templates
   that must hold (as per the static semantics in Fig.~\ref{fb-staticsem}).
 \item We solve the generated set of constraints using our fixpoint constraint
   solving algorithm \csolvestar, which reduces the constraint solving problem to
   an abduction problem in a Herbrand constraint system, and then relies on \csolve,
   our abduction solver for that domain. 
 \item We finally simplify the elaborated program by substituting the solution determined
   (for the region variables and predicate variables).
\end{enumerate}

@@

Function $\elabClassTable$ (Fig.~\ref{fig:fb-elabmeth}) elaborates
every definition in the class table $CT$, while accumulating
constraints. The constraints are finally solved by $\solve$
(\S\ref{sec:fb-constraintsolving}), which returns substitution
functions $\substFn_\rho$ and $\substFn_\varphi$ for free region and
predicate variables, respectively, introduced during templatization
and elaboration stages. The substitutions are applied to the class
table (and to the artifacts that make up the class table, recursively)
to compute a class table that maps classes to their fully
region-annotated definitions in $\FB$.

Note that if the original program in $\absof{\FB}$ contains unsafe
references, for example, a reference from a transferable region to a
stack regions, then the constraints generated during the elaboration
are not satisfiable. In such case, $\solve$ fails to solve
constraints, causing the program to be rejected.

>>>>

Fig.~\ref{fig:type-inference-algo} presents the high-level outline of the type
inference algorithm.
The algorithm consists of the following steps:
\begin{enumerate}
 \item The first step elaborates the input program by introducing \emph{(formal and actual) region parameters},
and \emph{region variables} (representing yet undetermined \emph{actual region parameters})
as follows:
\begin{enumerate}
\item For every class determine a set of \emph{region parameters} ($\rho$)
       it should be parametrized over.
\item For every field of the class, determine its region parametrized type by
       generating as many fresh \emph{actual region parameters} as required by its class (type),
       which in turn become formal region parameters of the containing class.
\item  For every method and function type, determine its corresponding region
   type template by introducing fresh region parameters as required by
   the type of each parameter and return value (in addition to the allocation region
   parameter).
 \item Elaborate every \C{new} expression, method call or function application by
   introducing \emph{fresh region variables} (as required by the region type
   template of the called method or function).
\item Introduce for each class and method, a \emph{predicate variable} ($\varphi$) to denote
   an undetermined set of constraints over the region parameters of that
   class/method.
\end{enumerate}
 \item In the second step, we analyze the program to generate a set of constraints that capture
   well-formedness requirements and subtyping relationships between type templates
   that must hold (as per the static semantics in Fig.~\ref{fb-staticsem}).
 \item We solve the generated set of constraints using our fixpoint constraint
   solving algorithm \csolvestar, which reduces the constraint solving problem to
   an abduction problem. 
% in a Herbrand constraint system, and then relies on \csolve,
% our abduction solver for that domain. 
 \item If the solver succeeds, it returns substitution functions $\substFn_\rho$ and
  $\substFn_\varphi$ for free region and predicate variables, respectively, introduced in the earlier
phases. We apply these substitutions to the elaborated program to produce the final program.
Note that if the original program in $\absof{\FB}$ contains unsafe
references, for example, a reference from a transferable region to a
stack regions, then the constraints generated during the elaboration
are not satisfiable. In such case, $\solve$ fails to solve
constraints, causing the program to be rejected.
\end{enumerate}

@@

Function $\elabClassTable$ (Fig.~\ref{fig:fb-elabmeth}) elaborates
every definition in the class table $CT$, while accumulating
constraints.
